A Practical Public Key Encryption Algorithm

### 2017-05-31 latex/secretkeybasedpublickeyencryption2/paper.tex
moved paper to new folder

\begin{abstract}
 We define a public-key encryption construction that is efficient in terms of key size, ciphertext size, key generation time, and encryption/decryption time. The construction requires only addition, multiplication, and modular reduction on large integers. The security of the scheme appears to rely only on the distribution of the random numbers generated and the correct selection of key parameters.
 \end{abstract}

@@

A fully homomorphic cipher supports multiple operations on ciphertexts, usually addition and multiplication. The public key encryption construction presented in this paper does not require a fully homomorphic cipher, only a a partially homomorphic cipher. We mention the term here only for the sake of completeness.

@@

\subsubsection{Basic Concept}
Points on (and off of) a lattice can be added together, and this effect is preserved even when a modular reduction step is applied, as long as the lattice basis and modulus are coprime and the point is smaller then the modulus. We exploit this modular reduction to both provide security and decrease the size of ciphertexts considerably.

\subsubsection{Storing Data In the Point Or In the Error}
When using a lattice to secretly store information, there are two places that the information can be stored: Either the information can be encoded as a point on the lattice, or the information can be encoded as the distance to the nearest lattice point.\\

Written in notation, a point on a lattice takes the form $p * q$, while a point plus some error takes the form $p * q + e$. In the former, the information is recovered by dividing by $p$ - or alternatively, multiplying by the modular multiplicative inverse of $p \mod n$. In the latter, the information is recovered by dividing by $p$ and taking the remainder. \\

We need to be able to encrypt 0, and using $p * q$ where $q = m$ is not suitable - therefore we must use $p * q + e$ where $e = m$. By itself, this cipher would be insecure; However, it does output a randomly distributed value. We use this randomly distributed value as a point on a secret lattice to produce the ciphertext.

\subsubsection{Encryption}
\begin{flalign*}
r_2 := (k_1 * r_1) + m\\
c := k_2 * r_2 \mod n
\end{flalign*}

\subsubsection{Decryption}
\begin{flalign*}
k_i := inverse(k_2, n)\\
r_2 := c * k_i \mod n\\
m := r_2 \mod k_1
\end{flalign*}

\subsubsection{Key Generation}
$k_1$ must be selected such that it is greater then the largest message that will be encrypted. If 256-bit messages are to be encrypted, then $k_1$ must be greater then 256 bits in size. We elect to utilize 264 bits for our $k_1$ values. \\

If the modulus is significantly larger then $k_2$, then $k_2$ could be determined by taking the gcd of $k_2 r_i$ pairs that happened to not reduce. This can be remedied by ensuring that $k_2$ is large enough. We use choose $k_2$ by generating a 256-bit number, and subtracting it from the modulus to obtain $k_2$. Selecting the point starting from $n$ instead of $0$ ensures that $k_2$ is large enough. If $k_2 >= n - k_2$, then gcd cannot be used to extract the value of $k_2$ even for consecutive small $q$.

\section{Security Analysis}
\subsection{Secret Key Cipher}
The security of $k * q \mod n$ as a cipher seems to reduce almost exclusively to the distribution of values for $q$, which is determined by $(k_1 * r_1) + m$. If $(k_1 * r_1) + m$ is considered to be a uniformly random value, then the secret key cipher is effectively equivalent to a One Time Pad, with a re-usable key. This is demonstrably not the case, as perfect secrecy is not provided by this cipher - A given ciphertext decrypts to one and only one particular plaintext message. We consider that $(k_1 * r_1) + m \mod n$  may not always follow a uniformly random distribution.

It appears that selecting $n$ as prime ensures a good distribution, regardless of the nature of the values of $k_1$ and $k_2$.\\

Encryptions of 0 take the form $p * k * q \mod n$. Given many such samples for random $q$, can p be recovered? If $x := p * k$, then the expressions look like $x * q \mod n$ for random $q$, which should be secure on the assumption that $q$ is unpredictable. If $q$ is unpredictable, then it should not be possible to recover $p$ from encryptions of $0$.\\

Encryptions of $m$ take the form $(p * k * q) + p * m \mod n$. Given many such samples for random q, can $p$ be recovered? If $p * k * q \equiv p * y \mod n$, and $y \mod n$ is unpredictable, then it should not be possible to recover $p$ from encryptions of $m$.\\

Under chosen ciphertext attack, we can obtain the value of $p_i \mod k$ by submitting the value of $1$ as a ciphertext to the decryption operation. Similarly, we can obtain the value of $p_i * c \mod n \mod k$ for some chosen values $c$. It is not yet clear how this information can be leveraged to recover $p$ or $p_i$.\\

If $k$ is a power of two, then $p_i \mod k$ directly reveals some of the lower bits of $p_i$. It is fantastically unlikely that $k$ will be a power of two if it is a randomly generated 264-bit number.

@@

\section{Metrics}
With the parameter sizes in this paper, a public key is 1600 bits in size, while a private key is about 1864 bits in size. Ciphertexts require 800 bits of space to encrypt 256-bit plaintext values. \\

With our public domain python implementation, we can encrypt 10,000 256-bit messages in approximately 0.58 seconds, while we can decrypt 10,000 ciphertexts in approximately 0.32 seconds. While a python implementation may not be the fastest possible, it performs well as much of the computation time is spent on BigNum operations anyways.\\

The public key encryption operation requires effectively the same amount of time as the secret key encryption operation; Consequently, there is little motivation to use the secret cipher as an encryption algorithm by itself. \\

The private key size could be decreased a small amount at the expense of computation time; The figure cited above represents the storage of $k_1, k_2, inverse(k_2, n)$. It is not necessary to store both $k_2$ and $inverse(k_2, n)$, as they can be calculated from each other. Considering the small amount of space it costs to store the inverse versus how long it takes to compute it, we choose to cache this value.

\subsection{Implementation Complexity}
The implementation complexity of the scheme is very low. The encryption and decryption functions for both constructions can be implemented with a single line of code each in the function body. Our implementation of the algorithms spans approximately 100 lines, including the documentation and comments.\\

@@

Assuming that BigNum operations can be implemented in a manner that does not leak timing information, it appears that a naive implementation of the algorithm is resistant to timing side channel analysis, as there are no tables or branches.

\subsection{Resource Asymmetry}
There is a computational asymmetry between encryption and decryption: Encryption takes longer then decryption. This is something that potentially could be tuned or exploited for other cryptographic constructions. Perhaps there is a way to distribute large public keys, which could be used as a part of a proof-of-space or proof-of-work scheme, or perhaps password hashing.

\section{Conclusion}
We define a conceptually simple, resource-practical public-key encryption algorithm. The security of our construction(s) appear to rest only on the distribution of intermediately generated random data and the correct selection of parameters.

\end{document}

>>>>

\begin{abstract}
 We define a public-key encryption construction that is efficient in terms of key size, ciphertext size, key generation time, and encryption/decryption time. The construction requires only addition, multiplication, and modular reduction on large integers.
 \end{abstract}

@@

A fully homomorphic cipher supports multiple operations on ciphertexts, usually addition and multiplication. The public key encryption construction presented in this paper does not require a fully homomorphic cipher, only a partially homomorphic cipher. We mention the term here only for the sake of completeness.

@@

\subsection{Secret Key Cipher}
\subsubsection{Basic Concept}
Points on (and off of) a lattice can be added together, and this effect is preserved even after being multiplied modulo a prime. We exploit this modular reduction to both provide security and decrease the size of ciphertexts considerably.

Multiple points can be superpositioned together via the addition operator, which provides a mechanism for randomizing the ciphertext. Given a point in the form of $(p * q_1) + (p ^ 2 * q_2)$, $p * q_1$ can be recovered by computing $(p * q_1) + (p ^ 2 * q_2) \mod p ^ 2$. This process can be repeated for successive powers of $p$. We utilize a three-tiered ciphertext using this technique, where the base point encodes the messages, and the squared/cubed points are both used for randomization.

\subsubsection{Key Generation}
\begin{flalign*}
p := random\_integer\\
k := random\_integer\\
k_i := inverse(k, n)
\end{flalign*}

\subsubsection{Key Generation}
$p$ must be selected such that it is greater then the largest message that will be encrypted. If 256-bit messages are to be encrypted, then $p$ must be greater then 256 bits in size. We elect to utilize 264 bits for our $p$ value. Similarly, if $k$ is to remain unguessable even in the face of an adversary with a quantum computer, then $k$ should be at least 256 bits in size. We choose $k$ as an offset from $n$, rather then as an offset from $0$ - this ensures that $k * q \mod n$ is always large enough to reduce, even if $q$ is small. This results in a larger value for $k$.\\

\subsubsection{Encryption}
\begin{flalign*}
t := (p * m) + (p ^ 2 * random\_integer) + (p ^ 3 * random\_integer)\\
c: = (k * t) \mod n
\end{flalign*}

\subsubsection{Decryption}
\begin{flalign*}
t := (k_i * c) \mod n\\
t := t \mod p ^ 3 \mod p ^ 2\\
m := t / p
\end{flalign*}

\section{Security Analysis}
\subsection{Secret Key Cipher}
The secret key cipher effectively consists of two separate sections; In the first section, a randomized ciphertext is created. The second section is a re-ordering of the ciphertext blocks via multiplication modulo n. This section exists to make the first section "harder" then it would be if it was not reduced modulo n. Note that $k$ is not sufficient to decrypt a given ciphertext. An adversary must acquire both $k$ and $p$, or find a way to recover $p$ without $k$. Given only $k$ and ciphertexts, it should still be difficult to recover a plaintext.\\

The security of $k * q \mod n$ as a cipher seems to reduce almost exclusively to the distribution of values for $q$, which is determined by $(p * m) + (p ^ 2 * r_0) + (p ^ 3 * r_1)$. If this was a uniformly random value, then the secret key cipher would be effectively equivalent to a One Time Pad, with a re-usable key. This is demonstrably not the case, as perfect secrecy is not provided by this cipher - A given ciphertext decrypts to one and only one particular plaintext message. We conclude that $(p * m) + (p ^ 2 * r_0) + (p ^ 3 * r_1)\mod n$  may not always follow a uniformly random distribution.\\

Encryptions of 0 take the form $x := k * ((p ^ 2 * r_0) + (p ^ 3 * r_1)) \mod n$. Given many such samples for random $r$, can $p$ or $k$ be recovered? Set $q := (p ^ 2 * r_0) + (p ^ 3 * r_1) \mod n$; If $q$ is unpredictable, then it should not be possible to recover $p$ or $k$ from encryptions of $0$.\\

Under chosen ciphertext attack, we can obtain the value of $(k_i \mod p ^ 3 \mod p ^ 2) / p$ by submitting the value of $1$ as a ciphertext to the decryption operation. Similarly, we can obtain the value of $(c * k_i \mod p ^ 3 \mod p ^ 2) / p$ for some chosen values $c$. It is not yet clear how this information can be leveraged to recover $p$ or $k_i$.\\

It is possible that somehow the relation between $p, p ^ 2, p ^3$ might be leveraged against the design. In such a situation, $p ^ 2$ and $p ^ 3$ can be changed to random numbers that are larger then $p ^ 2$ and $p ^ 3$, respectively. Currently, these points are generated as exponents of $p$ to save space, but this is not a requirement for the function of the construction.

@@

If necessary, the construction can be generalized to use larger quantities of randomization points. This would increase the size of the public key.

\section{Metrics}
With the parameter sizes in this paper, a public key is 2800 bits in size, while a private key is about 4400 bits in size. Ciphertexts require 1400 bits of space to encrypt 256-bit plaintext values. \\

With our public domain python implementation, we can encrypt with the public key 10,000 256-bit messages in approximately 0.66 seconds, while we can decrypt 10,000 ciphertexts in approximately 0.93 seconds. While a python implementation may not be the fastest possible, it performs well as much of the computation time is spent on BigNum operations anyways.\\

The private key size could be decreased at the expense of computation time; The figure cited above represents the storage of $p, p ^ 2, p ^ 3, k, inverse(k, n)$. It is not necessary to store $p ^ 2$, $p ^ 3$, or $inverse(k, n)$. Storing only $p$ and $k$ yields private key sizes of about 1000 bits. Considering the relatively small amount of space required, compared to how long it takes to compute a square/cube/modular inverse, we choose to cache these values.

\subsection{Implementation Complexity}
The implementation complexity of the scheme is very low. The encryption and decryption functions for the public key construction can be implemented with a single line of code each in the function body. Our implementation of the algorithms spans approximately 100 lines, including the documentation and comments.\\

@@

Assuming that BigNum operations can be implemented in a manner that does not leak timing information, it appears that a naive implementation of the algorithm is resistant to timing side channel analysis, as there are no tables or branches in the algorithm.

\section{Conclusion}
We define a conceptually simple, resource-practical public-key encryption algorithm. 
\end{document}
